---
title: "COVID-19 Vaccination Rates"
author: "Samantha Ruiz"
date: "3/3/2022"
output: 
  pdf_document:
    latex_engine: xelatex

---

#Import Vaccine Data
```{r}
vax <- read.csv( "~/Downloads/covid19vaccinesbyzipcode_test (1).csv" )
head(vax)
```

>Q1. What column details the total number of people fully vaccinated?

person_fully_vaccinated

>Q2. What column details the Zip code tabulation area?

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset?

"2022-03-01"

>Q4. What is the latest date in this dataset? 

```{r}

vax$as_of_date[nrow(vax)]
```

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

 18338 

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
round(18338/107604,2)*100
```



```{r}
library(lubridate)
```

```{r}
today() 
```

```{r}
age<- today() -ymd("2000-04-13")
age
```


```{r}
time_length(age, "year")
```

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```


>Q9. How many days have passed since the last update of the dataset?

```{r}
today()-vax$as_of_date[nrow(vax)]
```

>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length(unique(vax$as_of_date))
```



```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```
```{r}
reverse_zipcode(c('92037', "92109") )
```



Focus on the San Diego area

# Subset to San Diego county only areas 

```{r}
library(dplyr)
sd <- filter(vax, county == "San Diego")
dim(sd)
nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```



>Q11. How many distinct zip codes are listed for San Diego County?

```{r}
length(unique(sd$zip_code_tabulation_area))
```

>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd[which.max(sd$age12_plus_population),]
```


```{r}
sd$as_of_date[nrow(sd)]
```

let's do it wih the most recent data set
>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-02-22”? 

```{r}
sd.latest<- filter(sd, as_of_date=="2022-03-01")
mean(sd.latest$percent_of_population_fully_vaccinated, na.rm=TRUE)

```


```{r}
summary(sd.latest$percent_of_population_fully_vaccinated, na.rm=T)
```


>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-02-22”?



```{r}
hist(sd.latest$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)

ggplot(sd.latest) +
  aes(percent_of_population_fully_vaccinated)+ 
  geom_histogram()
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated)+
  geom_point()+
  geom_line()+
  theme_bw()
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")
```

>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
baseplot<- ggplot(ucsd) +
  aes(as_of_date, percent_of_population_fully_vaccinated)+
  geom_point()+
  geom_line()+
  theme_bw()


```


```{r}
baseplot+ 
  labs(title = "Vaccination Rate for CA 92037")
```

```{r}
vax.36 <- filter(vax, age5_plus_population>36144 & as_of_date =="2022-03-01")
head(vax.36)
```
```{r}
ave.36<- mean(vax.36$percent_of_population_fully_vaccinated, na.rm=T)
ave.36
```

```{r}
baseplot+geom_hline(yintercept= ave.36, linetype=2, color="red")
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated, na.rm=T)
```

>Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36)+
  aes(percent_of_population_fully_vaccinated)+
  geom_histogram()+
  xlim(c(0,1))
```

>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

```{r}

```


```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92037") %>%
  select(percent_of_population_fully_vaccinated)
```


>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)

ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="date", y="percent vaccinate",
       title="Vaccination RatesCross California",
       subtitle= "only areas with a population abouve 36k shown") +
  geom_hline(yintercept = ave.36, linetype=2)
```





